PANCHANGE_data_preprocessing

### 2020-08-05 PANCHANGE_data_preprocessing2020-07-15.Rmd
minor edits and joining respiratory to GLAD data

freq(coping.glad.respiratory.raw.numeric$respiratory.anosmia.binary)

@@

# Probable COVID-19 case identification
Algorithm from Menni et al. to identify probable COVID-19
Variables required:
- Sex: 	0 Female, 1 Male
- Other binary variables: 0 No, 1 Yes       #THESE NAMES WILL CHANGE WHEN IMPORTING DATA - FIND AND REPLACE
- "anosmia": loss of smell or taste  
- "pcough":	Persistent cough
- "sfatigue": Severe fatigue
- "smeals": Skipped meals
The below chunk needs to be re-thought - not as simple as categorising gender. 
```{r eval=FALSE, include=FALSE}
# Recode Gender to be binary for the model in RAMP
dat.respiratory.ramp <- dat.respiratory.ramp %>%
		  mutate(
		    gender_binary_ramp =
		      case_when(
		        Gender == "Male" ~ "Male",
		        Gender == "Female" ~ "Female",
		        Gender == !("Male" | "Female") ~ NA_real_
		      )
		  )
```

++KT: Needs to be developed further for RAMP dataset
```{r eval=FALSE, include=FALSE}
## COVID prediction model for COPING

## Formula to calculate probability 
dat.respiratory.coping <- dat.respiratory.coping %>%
		  mutate(
		    covid_probability = 
		      (-1.32) - 
		      (0.01*Age) + 
		      (0.44*Sex) + #we have sex here in COPING - use binary Gender variable in RAMP
		      (1.75*anosmia) + #will be renamed
		      (0.31*pcough) +  #will be renamed
		      (0.49*sfatigue) +  #will be renamed
		      (0.39*smeals) #will be renamed
		         )

## Calcualte the odds ratios 
dat.respiratory.coping <- dat.respiratory.coping %>%
		  mutate(
		    covid_odds_ratio = 
		      exp(covid_probability) / (1 + exp(covid_probability)
		                              )

## generate variable covid_odds_ratio_50 -> 0, 0,5 and 1 = 0, 1, 2 (that's a column). Previous code: egen covid_odds_ratio_50 = cut(covid_odds_ratio), at(0,0.5,1) icode
dat.respiratory.coping <- dat.respiratory.coping %>%
		  mutate(
		    covid_odds_ratio_50_numeric = 
		      case_when(
		        covid_odds_ratio >= 0 & covid_odds_ratio < 0.5 ~ "0", #cut off at 0.5 
		        covid_odds_ratio >= 0.5 & covid_odds_ratio <= 1 ~ "1"
		      )

#create factor with labels
dat.respiratory.coping <- dat.respiratory.coping %>%
  mutate(
    covid_odds_ratio_50 =
      recode_factor(
        covid_odds_ratio_50_numeric,
        "0" = "No COVID",
        "1" = "Probable COVID"
      )
  )

@@

descr(x = dat.respiratory.coping[,covid_variables],
      round.digits = 2,
      transpose = TRUE)

## looking at cases over 0.5
table(dat.respiratory.coping$covid_odds_ratio_50) 
```

>>>>

@@

Frequencies of binary COVID symptoms
```{r frequencies of COVID binary symptoms}
freq(coping.glad.respiratory.raw.numeric$respiratory.anosmia_binary)
freq(coping.glad.respiratory.raw.numeric$respiratory.pcough_binary)
freq(coping.glad.respiratory.raw.numeric$respiratory.fatigue_binary)
freq(coping.glad.respiratory.raw.numeric$respiratory.skipped_meals_binary)
```

Rename respiratory dataset joined to main dataset
```{r clean respiratory dataset}
coping.glad.respiratory <- coping.glad.respiratory.raw.numeric

dat.glad.coping <- left_join(
  x = dat.glad,
  y = coping.glad.respiratory,
  by = "ID"
)
```

# Probable COVID-19 case identification
Algorithm from Menni et al. to identify probable COVID-19
Variables required:
- Sex: 	0 Female, 1 Male
- Other binary variables: 0 No, 1 Yes       
- "respiratory.anosmia_binary": loss of smell or taste  
- "respiratory.pcough_binary":	Persistent cough
- "respiratory.fatigue_binary": Severe fatigue
- "respiratory.skipped_meals_binary": Skipped meals

++KT: Needs to be developed further for RAMP dataset
COVID prediction model for COPING
```{r eval=FALSE, include=FALSE}
# Formula to calculate probability 
dat.glad.coping <- dat.glad.coping %>%
		  mutate(
		    covid_probability = 
		      (-1.32) - 
		      (0.01*Age) + 
		      (0.44*Sex) + 
		      (1.75*respiratory.anosmia_binary) + 
		      (0.31*respiratory.pcough_binary) +  
		      (0.49*respiratory.fatigue_binary) +  
		      (0.39*respiratory.skipped_meals_binary) 
		         )

# Calcualte the odds ratios 
dat.glad.coping <- dat.glad.coping %>%
		  mutate(
		    covid_odds_ratio = 
		      exp(covid_probability) / (1 + exp(covid_probability))

# generate variable covid_odds_ratio_50 -> 0, 0,5 and 1 = 0, 1, 2 (that's a column)
dat.glad.coping <- dat.glad.coping %>%
		  mutate(
		    covid_odds_ratio_50_numeric = 
		      case_when(
		        covid_odds_ratio >= 0 & covid_odds_ratio < 0.5 ~ "0", #cut off at 0.5 
		        covid_odds_ratio >= 0.5 & covid_odds_ratio <= 1 ~ "1"
		      )

#create factor with labels
dat.glad.coping <- dat.glad.coping %>%
  mutate(
    covid_odds_ratio_50 =
      recode_factor(
        covid_odds_ratio_50_numeric,
        "0" = "No COVID",
        "1" = "Probable COVID"
      )
  )

@@

descr(x = dat.glad.coping[,covid_variables],
      round.digits = 2,
      transpose = TRUE)

## looking at cases over 0.5
table(dat.glad.coping$covid_odds_ratio_50) 
```
